Synpotic Soil Characterization

### 2024-10-23 3-reports/manuscript_figures.Rmd
Fe/P plotting

>>>>

```{r analytesx-3, fig.height = 10, fig.width = 8}

plot_transect_as_x = function(data, YLAB = "", TITLE = "", SUBTITLE = "", SCALES = "free_x"){

  data = data %>% mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
    filter(!is.na(value))

  fit_lme_hsd_transect = function(dat){
    # this function will assign post-hoc HSD letters to the transect
    # use lme4::lmer -> multcomp::glht -> multcomp::cld
    library(multcomp)

    lmer = lme4::lmer(value ~ transect + (1|site), data = dat)
    h = summary(glht(lmer, linfct = mcp(transect = "Tukey")), test = adjusted("holm"))
    x = cld(h, decreasing = TRUE) # decreasing order of HSD letters
    x$mcletters$monospacedLetters %>% # sigh
      as.data.frame() %>% 
      rownames_to_column("transect") %>% 
      rename(groups = ".") %>% 
      mutate(groups = str_remove_all(groups, " "),
             groups = toupper(groups))

  # make summary table for HSD letters ----
  hsd_transect_overall = 
    data %>% 
    group_by(region) %>% 
    do(fit_lme_hsd_transect(.))

  ## then calculate the y-axis position for the labels
  lme_y =
    data %>% 
    #group_by(region) %>% 
    dplyr::summarise(max = max(value, na.rm =T),
                     y = max + max/5) %>% 
    ungroup()

  ## combine them
  lme_for_graphs = 
    hsd_transect_overall %>% 
    cross_join(lme_y %>% dplyr::select(-max))

  #
  # graphs ----

  ##  ggplot(data,
  ##         aes(x = transect,
  ##             y = value
  ##             #shape = horizon,
  ##             #alpha = horizon
  ##         )) +
  ##    # plot points
  ##    geom_point(aes(group = site,
  ##                   color = site),
  ##               size = 2.5, stroke = 1,
  ##               position = position_dodge(width = 0.4))+
  ##    # plot HSD letters
  ##    geom_text(data = lme_for_graphs, aes(y = y, label = groups), size = 5)+
  ##    scale_alpha_manual(values = c(1, 0.3))+
  ##    scale_color_manual(breaks = c("CRC", "PTR", "OWC", "GCW", "MSM", "GWI"), values = c('#ED6e85', '#7f4420', '#ffc115', '#90BE6D', '#03045E', '#00B4D8'))+
  ##    facet_wrap(~region, 
  ##               labeller = as_labeller(c("CB" = "Chesapeake Bay", "WLE" = "Lake Erie")))+
  ##    theme_kp()+
  ##    labs(x = "",
  ##        # y = YLAB,
  ##        # title = TITLE,
  ##        # subtitle = SUBTITLE,
  ##         shape = "Site",
  ##         alpha = "Horizon",
  ##         color = "Site")+
  ##    theme(legend.position = "right", legend.box = "vertical")+
  ##    NULL

  ggplot(data,
         aes(x = transect,
             y = value
             #shape = horizon,
             #alpha = horizon
         )) +
    # plot points

    geom_violin(aes(fill = transect), alpha = 0.4)+
    geom_jitter(width = 0.1)+
    geom_text(data = lme_for_graphs, aes(y = y, label = groups), size = 5)+
    labs(x = "",
         y = "Extractable P, μg/g")+
    facet_wrap(~region)+
    scale_fill_manual(values = pal_transect)+
    labs(x = "",
         y = YLAB,
         title = TITLE,
         subtitle = SUBTITLE,
         shape = "Site",
         alpha = "Horizon",
         color = "Site")+
    theme(legend.position = "none")+
    NULL

gg_fe = plot_transect_as_x(data_combined_wide %>% rename(value = Fe),
                   YLAB = "Extractable Fe, μg/g")

gg_p = plot_transect_as_x(data_combined_wide %>% rename(value = P),
                   YLAB = "Extractable P, μg/g")

gg_weoc = plot_transect_as_x(data_combined_wide %>% rename(value = WEOC),
                   YLAB = "Water-Extractable Organic C, μg/g")

cowplot::plot_grid(gg_fe, gg_p, ncol = 1)
